Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-05-09 manuscript/manuscript.tex
Changes in manuscript

\begin{figure*}
\centering
\includegraphics[width=0.9\linewidth]{figures/linear-D.pdf}
\caption{
    Differences between the gravity fields generated by the numerical 
    model and the analytical solution for (a) a thin spherical shell of 
    1km of thickness and (b) a thick spherical shell of 35km of thickness,
    both with the linear density defined on equation 
    \ref{eq:density-linear}.
    The computations were performed on the four grids described in 
    Table \ref{tab:grids}, for different values of the distance-size 
    ratio $D$.
    Due to the linearity of the density function the density-based 
    discretization algorithm is not applied.
    If the difference is less than 0.1\%, we consider that the model 
    has achieved an acceptable accuracy for the corresponding value of 
    $D$.
    }
\label{fig:D-linear}
\end{figure*}

@@

On Figure \todo[inline]{referencia a figura grid search} we show 
the resulting percentage difference for each gravitational field.
The points inside the dotted lines are the ones that present a 
numerical error lower than the 0.1\% threshold, i.e. they represent the 
eligible ($D$, $\delta$) pairs.
The one that produces the more efficient computation must be chosen as 
the default $D$ and $\delta$ value.

@@

On the other hand, we have no a priori information about the numerical 
error of points with $\delta > \delta_p$ (region C on 
Figure \ref{fig:D-delta-grid-search}).
Nevertheless, the numerical error exploration shown on
Figure \todo[inline]{ref a figura grid search} doesn't present any 
point in this region with error lower than the threshold.
Even if we are analysing a single case, it proves that no point in 
region C is eligible as default ($D$, $\delta$) pair because it should 
ensure numerical accuracy for every exponential density scenario.

@@

In conclusion, the ($D$, $\delta$) point that minimises the number of 
discretizations while keeping an acceptable numerical error under the 
0.1\% threshold is the one on which the ratio $D$ is equal to 
$D_\text{linear}$ and the delta ratio is $\delta_p$.

@@

On Figures \ref{fig:delta-exponential}(a) and 
\ref{fig:delta-exponential}(b) we show the density functions used in 
the comparison and the resulting percentage difference for each 
gravitational field.
The former shows the results for the thin spherical shell, while the 
later shows the ones corresponding to the thick shell.
From these results we can conclude that for the potential and $g_z$ 
computation, a $\delta = 0.3$ is needed in order to achieve an 
acceptable numerical error, while a $\delta = 0.02$ is required for the 
$g_{zz}$ Marussi tensor component.

>>>>

\begin{figure*}
\centering
\includegraphics[width=0.9\linewidth]{figures/linear-D.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model 
    and the analytical solution for a thin and a thick spherical shell of 1km 
    and 35km of thickness, respectively.
    Both models have the linear density defined on equation 
    \ref{eq:density-linear}.
    The computations were performed on the four grids described in 
    Table \ref{tab:grids}, for different values of the distance-size 
    ratio $D$.
    Due to the linearity of the density function the density-based 
    discretization algorithm is not applied.
    If the difference is less than 0.1\%, we consider that the model 
    has achieved an acceptable accuracy for the corresponding value of 
    $D$.
    }
\label{fig:D-linear}
\end{figure*}

@@

On Figure \ref{fig:grid-search} we show 
the resulting percentage difference for each gravitational field.
The points inside the dotted lines are the ones that present a 
numerical error lower than the 0.1\% threshold, i.e. they represent the 
eligible ($D$, $\delta$) pairs.
The one that produces the more efficient computation must be chosen as 
the default $D$ and $\delta$ value.

@@

On the other hand, we have no a priori information about the numerical 
error of points with $\delta > \delta_p$ (region C on 
Figure \ref{fig:D-delta-grid-search}).
Nevertheless, the numerical error exploration shown on
Figure \ref{fig:grid-search} doesn't present any 
point in this region with error lower than the threshold.
Even though we are analysing a single case, it proves that no point in 
region C is eligible as default ($D$, $\delta$) pair because it should 
ensure numerical accuracy for every exponential density scenario.

@@

In conclusion, the ($D$, $\delta$) point that minimises the number of 
discretizations while keeping an acceptable numerical error under the 
0.1\% threshold is the one whose ratio $D$ is equal to 
$D_\text{linear}$ and its delta ratio is $\delta_p$.

In order to determine a default value of $\delta$ that ensures an 
acceptable accuracy for every exponential density case we have 
performed comparisons between the numerical model and the analytical 
solutions, keeping the distance-size ratio $D$ equal to the values 
obtained for the linear density case while exploring different values of 
$\delta$.

We have considered the same spherical shell models we have used in the 
linear density case: a thin one (1km of thickness) and a thick one 
(35km of thickness) composed by $30^\circ \times 30^\circ$ tesseroids.
But in this case, we assigned an exponential density function like the 
one shown in equation \ref{eq:density-exp} to each shell, choosing 
different values of $b$ in order to explore distinct density 
variations.

Finally, we have computed the gravitational potential ($V$), the 
vertical component of its gradient ($g_z$) and the $g_zz$ component of 
the Marussi tensor on the big grid (see Table \ref{tab:grids}) and 
calculated the percentage difference between these numerical results and 
their corresponding analytical solutions.
Our intention is to set the default $\delta$ as the highest value whose 
numerical error is bellow the $0.1\%$ threshold.

@@

On Figure \ref{fig:delta-exponential} we show the density functions 
used on each comparison and the resulting percentage difference for each 
gravitational field.
The first column shows the results for the thin spherical shell, while the 
second one shows the ones corresponding to the thick shell.
From these we can conclude that a $\delta = 0.2$ is needed in order to 
achieve an acceptable numerical error for the computation of the 
potential and the $g_z$ component.
Because the same comparison using the other components of the gradient ($g_x$ 
and $g_y$) could not be carried out, we will extrapolate this result 
setting a default $\delta = 0.2$ for their computation.

On the other hand, the comparison with the $g_{zz}$ component of the 
Tensor shows a difference between the thin and the thick spherical 
shell: the former needs a $\delta = 0.2$, but the latter requires a 
$\delta = 0.01$.
In order to guarantee a good accuracy for every case, we will keep the 
more conservative value of $\delta = 0.02$ for the computation of the 
Marussi tensor components.
